Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix some typos

In this work,
we further develop the research on pulverised systems by providing two key contributions:
%
\begin{enumerate}
 \item we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
 \item we accordigly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
\end{enumerate}
%
To evaluate \ourframework{}, we exercise it through two case studies:
a small-scale real-wold demonstration, and a city-scale simulated scenario.

@@

The \emph{deployment view} is a well-known architectural viewpoint for software systems,
concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
 and supported by modelling notations like \emph{UML Deployment diagrams}.
%
Here, we briefly introduce deployment and reconfiguration based on the conceptual characterisation of~\cite{DBLP:journals/jss/ArcangeliBL15,carzaniga1998characterization}.
%
A \emph{site} is a set of computers (\emph{hosts}) that may host a \emph{software system}, i.e., a product described by a coherent collection of artefacts and consisting of a set of \emph{deployment units} that can be independently operated.
%
\emph{(Software) Deployment} is the process of moving and making a software system available and operational from one or more \emph{producer sites} to a target set of \emph{consumer sites}, also called the \emph{deployment domain}.
%
The \emph{deployment plan} defines 
 the mapping between the software system
 and the deployment domain,
 possibly augmented with further information (e.g., metadata, constraints, preferences).
%
Common deployment-related activities include:
(i) \emph{release/update} of the software system at the producer site;
(ii) \emph{installation/uninstallation} at/from the consumer sites;
(iii) \emph{activation/deactivation}, for starting/stopping the components;
(iv) \emph{reorganisation} of the software system;
and (v) \emph{redistribution}, i.e., changing the deployment plan.
%
According to the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
  the problem of (automatic) deployment of distributed software systems can be addressed by considering
 (i) the nature of the software to be deployed (e.g., how the software system is split into components);
 (ii) the nature of the deployment domain (i.e. the characteristics and topology of the available infrastructure);
 (iii) how the deployment is designed (e.g., how the deployment plan is specified);
 and 
 (iv) how the deployment is performed (e.g., how deployment activities are carried out).
%
From an operational point of view,
 deployment may be supported by a so-called \emph{runtime} or \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.
%
Related work covering these issues is provided in \Cref{sec:rw}.
%
In the following, we introduce the deployment approach 
  that we extend and upon which we develop \ourframework{}.

@@

A framework/runtime with configuration files enforces declarativity at the configuration level,
but its flexibility is limited,
as options not accounted for at design time can hardly be injected through configuration files.
%
On the other hand,
a (well-designed) library can be flexible as the host language can be used to express peculiar configurations
and is easier to adopt for users acquainted with the host language.
%
However,
its configuration may quickly become imperative and more challenging to maintain as complexity grows.
%
Finally, a \ac{dsl}~\cite{dsl-book-voelter} can be designed to be as expressive as needed,
but it has high maintenance cost
(as the language maintenance stacks upon the library/\ac{api}) and, potentially,
a steep learning curve due to the need to learn a new (custom) language.

@@

\begin{lstlisting}[
    float=ht,
    language=pulv,
    caption={
        Example of \dslSys{} usage.
        This code snippet defines three capabilities and a logical system composed of two device types,
        whose components are bound to the set of capabilities they needs to execute.
    },
    label=lst:system-configuration,
    escapechar=\%,
    numbers=left
]
object HighCPU : Capability%\label{capab:start}%
object LowLatencyComm : Capability
object EmbeddedDevice : Capability%\label{capab:end}%

@@

We assume that the power consumption of the device is linearly dependent on the CPU usage,
and we thus decided to estimate the \ac{epi}~\cite{DBLP:conf/islped/ShaoB13}
of a typical smartphone ($EPI_{device}$) and server ($EPI_{cloud}$) CPUs.
%
To do so, we take the \ac{tdp} of a
Qualcomm Snapdragon 888 ($5W$)
and an
Intel Xeon Platinum P-8124 ($220W$),
and we divide it by the score the CPU obtained in the popular Passmark benchmark\footnote{\url{https://www.cpubenchmark.net/}}
($9362$ and $22674$, respectively)
obtaining an indication of the power per benchmark point
($\nicefrac{5W}{9362pts}$ and \nicefrac{220W}{22674pts} respectively).
%
We then used these ratios to estimate the relative \ac{epi} of the two CPUs,
obtaining an $EPI_{ratio}$ close to 1:18.
%
Thus,
we take the \ac{epi} estimation for the server CPU from~\cite{DBLP:conf/islped/ShaoB13},
(approximately 10J per instruction, considering 36 logical cores of our reference processor)
and we assume the \ac{epi} of the smartphone CPU to be 18 times lower.
%
Once the \acp{epi} of the two platforms are known,
we estimate the number of instructions of the pulverised components as follows:
we assume the application to be a heavy-duty game,
capable of draining the battery of a smartphone
(approximately 4500mAh in capacity, and thus, at 3.3V, 14.85Wh, 53460J)
in 6 hours of continuous usage,
which divided by the \ac{epi} of the smartphone CPU gives us the number of instructions.
%
Additionally,
we account for a variable quota of instructions that captures other activities of the devices
(screen, operating system, other applications, etc.).

@@

\subsubsection{Implementation}
We implemented our simulations by interfacing \ourframework{}
with the Alchemist Simulator~\cite{PianiniJOS2013}
through a dedicated platform module.
%
We ran experiments varying the number of participants (hence, smartphones): 100, 300, and 1000
and the \ourframework{} reconfiguration rules:
$\Uparrow$ (behaviour always in the smartphones),
$\Updownarrow^{60}_{20}$,
$\Updownarrow^{80}_{20}$,
$\Updownarrow^{60}_{40}$,
$\Updownarrow^{80}_{40}$,
$\Downarrow$ (behaviour always in the cloud).
%
All the hybrid configurations
(every configuration except $\Uparrow$ and $\Downarrow$)
require runtime reconfiguration.
%
We run 10 repetitions or each experiment with a different random seed;
all results are averaged over them.
%
The experiment has been open-sourced,
and the code has been archived on Zenodo~\cite{https://doi.org/10.5281/zenodo.7948801}.
%
\begin{figure}
    \includegraphics[width=\linewidth]{figures/simulation-screenshot.png}
    \caption{
        A snapshot of the large-scale simulation.
        Smartphones are represented as coloured circles,
        with the colour indicating their battery level
        (green when charged, red when low).
    }
    \label{fig:simulation-screenshot}
\end{figure}
%
\Cref{fig:simulation-screenshot} shows a snapshot of the simulation

\subsubsection{Results}
%
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
    \caption{
        Simulation results for 100 (top), 300 (centre), and 1000 (bottom) devices;
        showing all four metrics:
        $P_{cloud}$ ($kW$) in green,
        $P_{devices}$ ($kW$) in blue,
        Distance in yellow, $\$_{cloud}$ in dark purple.
        %
        $P_{cloud}$ and $P_{devices}$ are stacked to show the overall power consumption.
        %
        Different reconfiguration rules are indicated on the horizontal axis.
        %
        Runtime reconfiguration of the deployment opens more opportunities to find the
        cost, energy use, and \ac{qos} trade-off that best fits the current context.
    }
    \label{fig:results}
\end{figure}
%
Results are summarised in \Cref{fig:results}.
%
$\Downarrow$ has the worst \ac{qos}
but also the lowest cloud cost and overall energy consumption.
%
Opposedly, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
%
The ability to reconfigure the deployment at runtime
by moving the behaviour from the cloud to the smartphones and vice versa
depending on the battery level allows us to find different trade-offs between these two extremes.

>>>>

In this work,
we further develop the research on pulverised systems by providing two key contributions:
%
\begin{enumerate}
 \item we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
 \item we accordigly provide a practical implementation of pulverisation (with reconfiguration rules) called \emph{\ourframework{}}, developed in the Kotlin programming language.
\end{enumerate}
%
To evaluate \ourframework{}, we exercise it through two case studies:
a small-scale real-world demonstration, and a city-scale simulated scenario.

@@

The \emph{deployment view} is a well-known architectural viewpoint for software systems,
concerned with the mapping of \emph{software components} to \emph{physical machines}~\cite{DBLP:journals/software/Kruchten95}
 and supported by modeling notations like \emph{UML Deployment diagrams}.
%
Here, we briefly introduce deployment and reconfiguration based on the conceptual characterisation of~\cite{DBLP:journals/jss/ArcangeliBL15,carzaniga1998characterization}.
%
A \emph{site} is a set of computers (\emph{hosts}) that may host a \emph{software system}, i.e., a product described by a coherent collection of artefacts and consisting of a set of \emph{deployment units} that can be independently operated.
%
\emph{(Software) Deployment} is the process of moving and making a software system available and operational from one or more \emph{producer sites} to a target set of \emph{consumer sites}, also called the \emph{deployment domain}.
%
The \emph{deployment plan} defines 
 the mapping between the software system
 and the deployment domain,
 possibly augmented with further information (e.g., metadata, constraints, preferences).
%
Common deployment-related activities include:
(i) \emph{release/update} of the software system at the producer site;
(ii) \emph{installation/uninstallation} at/from the consumer sites;
(iii) \emph{activation/deactivation}, for starting/stopping the components;
(iv) \emph{reorganisation} of the software system;
and (v) \emph{redistribution}, i.e., changing the deployment plan.
%
According to the analytical framework of~\cite{DBLP:journals/jss/ArcangeliBL15},
  the problem of (automatic) deployment of distributed software systems can be addressed by considering
 (i) the nature of the software to be deployed (e.g., how the software system is split into components);
 (ii) the nature of the deployment domain (i.e. the characteristics and topology of the available infrastructure);
 (iii) how the deployment is designed (e.g., how the deployment plan is specified);
 and 
 (iv) how the deployment is performed (e.g., how deployment activities are carried out).
%
From an operational point of view,
 deployment may be supported by a so-called \emph{runtime} or \emph{middleware}~\cite{DBLP:journals/cacm/GazisK22}.
%
Related work covering these issues is provided in \Cref{sec:rw}.
%
In the following, we introduce the deployment approach 
  that we extend and upon which we develop \ourframework{}.

@@

A framework/runtime with configuration files enforces declarativity at the configuration level,
but its flexibility is limited,
as options not accounted for at design time can hardly be injected through configuration files.
%
On the other hand,
a (well-designed) library can be flexible as the host language can be used to express peculiar configurations
and is easier to adopt for users acquainted with the host language.
%
However,
its configuration may quickly become imperative and more challenging to maintain as complexity grows.
%
Finally, a \ac{dsl}~\cite{dsl-book-voelter} can be designed to be as expressive as needed,
but it has a high maintenance cost
(as the language maintenance stacks upon the library/\ac{api}) and, potentially,
a steep learning curve due to the need to learn a new (custom) language.

@@

\begin{lstlisting}[
    float=ht,
    language=pulv,
    caption={
        Example of \dslSys{} usage.
        This code snippet defines three capabilities and a logical system composed of two device types,
        whose components are bound to the set of capabilities they need to execute.
    },
    label=lst:system-configuration,
    escapechar=\%,
    numbers=left
]
object HighCPU : Capability%\label{capab:start}%
object LowLatencyComm : Capability
object EmbeddedDevice : Capability%\label{capab:end}%

@@

We assume that the power consumption of the device is linearly dependent on the CPU usage,
and we thus decided to estimate the \ac{epi}~\cite{DBLP:conf/islped/ShaoB13}
of a typical smartphone ($EPI_{device}$) and server ($EPI_{cloud}$) CPUs.
%
To do so, we take the \ac{tdp} of a
Qualcomm Snapdragon 888 ($5W$)
and an
Intel Xeon Platinum P-8124 ($220W$),
and we divide it by the score the CPU obtained in the popular Passmark benchmark\footnote{\url{https://www.cpubenchmark.net/}}
($9362$ and $22674$, respectively)
obtaining an indication of the power per benchmark point
($\nicefrac{5W}{9362pts}$ and \nicefrac{220W}{22674pts} respectively).
%
We then used these ratios to estimate the relative \ac{epi} of the two CPUs,
obtaining an $EPI_{ratio}$ close to 1:18.
%
Thus,
we take the \ac{epi} estimation for the server CPU from~\cite{DBLP:conf/islped/ShaoB13},
(approximately 10J per instruction, considering 36 logical cores of our reference processor)
and we assume the \ac{epi} of the smartphone CPU to be 18 times lower.
%
Once the \acp{epi} of the two platforms are known,
we estimate the number of instructions for the pulverised components as follows:
we assume the application to be a heavy-duty game,
capable of draining the battery of a smartphone
(approximately 4500mAh in capacity, and thus, at 3.3V, 14.85Wh, 53460J)
in 6 hours of continuous usage,
which divided by the \ac{epi} of the smartphone CPU gives us the number of instructions.
%
Additionally,
we account for a variable quota of instructions that captures other activities of the devices
(screen, operating system, other applications, etc.).

@@

\subsubsection{Implementation}
We implemented our simulations by interfacing \ourframework{}
with the Alchemist Simulator~\cite{PianiniJOS2013}
through a dedicated platform module.
%
We ran experiments varying the number of participants (hence, smartphones): 100, 300, and 1000
and the \ourframework{} reconfiguration rules:
$\Uparrow$ (behaviour always in the smartphones),
$\Updownarrow^{60}_{20}$,
$\Updownarrow^{80}_{20}$,
$\Updownarrow^{60}_{40}$,
$\Updownarrow^{80}_{40}$,
$\Downarrow$ (behaviour always in the cloud).
%
All the hybrid configurations
(every configuration except $\Uparrow$ and $\Downarrow$)
require runtime reconfiguration.
%
We run 10 repetitions of each experiment with a different random seed;
all results are averaged over them.
%
The experiment has been open-sourced,
and the code has been archived on Zenodo~\cite{https://doi.org/10.5281/zenodo.7948801}.
%
\begin{figure}
    \includegraphics[width=\linewidth]{figures/simulation-screenshot.png}
    \caption{
        A snapshot of the large-scale simulation.
        Smartphones are represented as coloured circles,
        with the colour indicating their battery level
        (green when charged, red when low).
    }
    \label{fig:simulation-screenshot}
\end{figure}
%
\Cref{fig:simulation-screenshot} shows a snapshot of the simulation

\subsubsection{Results}
%
\begin{figure}
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=100.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=300.0.pdf}\\
    \includegraphics[width=\columnwidth]{figures/cloud_cost-device_consumption-cloud_consumption-distance-device=1000.0.pdf}
    \caption{
        Simulation results for 100 (top), 300 (center), and 1000 (bottom) devices;
        showing all four metrics:
        $P_{cloud}$ ($kW$) in green,
        $P_{devices}$ ($kW$) in blue,
        Distance in yellow, $\$_{cloud}$ in dark purple.
        %
        $P_{cloud}$ and $P_{devices}$ are stacked to show the overall power consumption.
        %
        Different reconfiguration rules are indicated on the horizontal axis.
        %
        Runtime reconfiguration of the deployment opens more opportunities to find the
        cost, energy use, and \ac{qos} trade-off that best fits the current context.
    }
    \label{fig:results}
\end{figure}
%
Results are summarised in \Cref{fig:results}.
%
$\Downarrow$ has the worst \ac{qos}
but also the lowest cloud cost and overall energy consumption.
%
Opposedly, $\Uparrow$ improves the \ac{qos}, but costs and energy use grow significantly.
%
The ability to reconfigure the deployment at runtime
by moving the behaviour from the cloud to the smartphones and vice versa
depending on the battery level allows us to find different trade-offs between these two extremes.
